\chapter*{Declaration on the Use of Generative Artificial Intelligence}

I declare that the following generative AI tools were used during the preparation of this work in the ways stated below:

\begin{itemize}
  \item \textbf{ChatGPT (OpenAI)} was used in the period \textbf{August 2024--April 2026}.
  It was used for brainstorming and refining the scope and structure, and for AI-assisted drafting and iterative revision of written text across substantial parts of the work. It was also used for discussion of implementation options, technical trade-offs, and explanatory framing. All generated outputs were critically reviewed, edited, and integrated by the author.

  \item \textbf{Codex (OpenAI)} was used in the period \textbf{January 2026--April 2026}.
  It was used for comprehensive review and targeted rewriting of selected passages with broader document context taken into account. The author reviewed, adapted, and incorporated all resulting outputs.

  \item \textbf{GitHub Copilot (GitHub/Microsoft)} was used in the period \textbf{August 2025--April 2026}.
  It was used primarily as code completion and assistance with repetitive implementation patterns. The final code and all technical decisions were authored, reviewed, and validated by the author.

  \item \textbf{Cursor (AI-assisted IDE)} was used in the period \textbf{December 2025--April 2026}.
  It was used to accelerate refactoring and repetitive edits (e.g., propagating interface changes across modules and updating related boilerplate). The author reviewed, corrected, and integrated all resulting changes.
\end{itemize}

After using the tools listed above, I thoroughly reviewed and modified the outputs as needed and I fully accept responsibility for the final content of this work, including all technical claims, experimental results, and interpretations.
