\chapter{Related Work} \label{chapter:ch1_related_work}

This thesis explores the concept of \emph{in-workspace spatial authoring}, where the user specifies spatial intent directly in the real workspace using AR-inspired input, supported by external sensing. The system then expands this intent into the required motion and tool commands for the robot. The goal of this chapter is not an exhaustive survey; rather, it provides a focused overview of
\begin{enumerate*}[label=(\roman*)]
    \item what directions have been explored in the literature for reducing friction in robot programming, and
    \item where our approach fits in the landscape of related work.
\end{enumerate*}
Together with the requirement derivation and design comparisons in \autoref{chapter:ch2_analysis}, this chapter forms the survey-and-selection basis used to justify the implemented system direction.

\section[AR/VR interfaces: context from surveys]{AR/VR and human-facing interfaces for robot programming: context from surveys}

A common theme in the robot programming literature is that robot deployment effort is often dominated by programming and integration rather than pure execution time. In high-mix low-volume scenarios, this becomes especially costly: an IFR industry perspective \cite{IFR2023TraditionalVsNew} notes that programming and integration can account for 50--70\,\% of the cost of a robot application. Surveys on human-robot interaction for programming \cite{VILLANI201866} similarly motivate alternative programming interfaces for improving usability and productivity. AR/VR have become a common interface direction for both offline and online programming \cite{Makhataeva2020, DIANATFAR2021407}.

Across these works, a distinction emerges between \emph{visualization}-focused (showing robot state, planned motion, or safety zones in context) and \emph{authoring}-focused (letting the user create or edit robot behavior directly in the workspace) approaches. For our thesis, we focus on the latter: authoring requires not only visualization but also robust spatial input, registration to world or robot coordinates, and a program representation that is simple enough for non-expert users yet expressive enough for useful tasks. 

\section{AR robot programming systems}

Existing AR robot programming systems span a wide design space. We summarize representative systems using the following two dimensions:
\begin{enumerate}[label=(\roman*)]
    \item \emph{where authoring happens} (online in the real workspace vs.\ offline in a simulated environment/digital twin), and
    \item \emph{how motion is specified} (point-by-point waypoints, teach-by-demonstration trajectories, or higher-level rules).
\end{enumerate}

\subsection{Online trajectory authoring} \label{subsec:related_work_online_trajectory_authoring}

Ong et al.~\cite{Ong2020} present \emph{ARRPS}, an augmented-reality-based system for intuitive robot programming that transforms the robot workcell into an AR environment and uses a handheld pointer to define 3D points and paths for the robot to follow. Their system addresses feasibility by integrating motion planning, collision checking, and plan validation to reduce the amount of robot-programming expertise needed during authoring.

This line of work is close in spirit to our approach, since the user also operates in the real scene and the system is responsible for turning spatial input into feasible robot motion. However, the interaction is primarily waypoint/path-centric --- the user specifies \emph{how} the robot should move rather than stating higher-level, task-oriented intent.

\subsection{Online rule-based authoring}

Ikeda and Szafir~\cite{Ikeda2024} propose \emph{PRogramAR}, an AR system for end-user robot programming where, instead of explicit low-level motion specifications, the user creates \emph{trigger-action} rules (if-then, while-do). Users define spatial regions and reference tracked objects to express conditions that trigger robot actions. This allows users to create program logic without needing to specify low-level motion details.

This direction is closer to our focus because it aligns with the motivation for task-oriented use cases --- the user should not be forced to manually create the full internal sequence of approach, action, and retreat steps if the system can infer them from intent and scene context. However, \emph{PRogramAR} emphasizes reactive, condition-triggered actions and does not target continuous trajectories as a primary interaction primitive. Our approach instead focuses on fixed task-oriented use cases: if a repetitive task family is not yet supported, we add a new use case that implements the required interpretation logic on top of shared sensing/authoring/execution infrastructure.

\subsection{Offline trajectory authoring}

Both of the approaches above are online, meaning that the user is present in the real workspace while authoring. This can interrupt production (robot availability) and introduces safety considerations during test runs. An alternative is therefore offline authoring, which mirrors traditional offline programming: a virtual robot is programmed in a simulated or AR environment and later transferred to the real robot.

Müller et al.~\cite{Muller2024} propose an offline programming workflow inspired by kinesthetic teaching, but performed through a virtual representation with tracked input, followed by transfer to a real robot via calibration and export. The paper emphasizes practical transfer steps and reports a user study that measures programming-time reductions relative to a baseline. As with offline programming in general, the approach trades reduced robot downtime for added setup and transfer complexity, and the achieved accuracy depends on calibration quality.

The approach in our thesis targets a different trade-off --- we focus on online authoring through prepared task-oriented use cases in order to minimize the required expertise and time needed to create and execute programs for frequently changing tasks. This also avoids offline scene replication/transfer effort and the associated calibration pipeline.

\subsection{Online point-by-point hand-tracking authoring}

Hand tracking is an attractive input modality for robot programming because it is intuitive and does not require handheld devices. However, practical systems often face constraints in accuracy and occlusion robustness. Vysocký et al.~\cite{Vysocky2023} implement a hand-tracking-based system for programming robot motions and subsequently executing the created program.

Robot motion is programmed by defining waypoints in the workspace using hand gestures. Hand tracking is accomplished using several calibrated RGB-D cameras with a MediaPipe-based \cite{MediaPipeRoot} hand landmark detector. Their experiment with 20 participants shows that gesture-based waypoint creation can be faster on average than conventional methods (teach pendant jogging and hand-guiding), but that waypoint accuracy is lower for hand-tracking. This provides a useful data point for the input trade-off in our thesis: if we require precise pose specification in world/robot coordinates, we may have to consider other input modalities.

Similarly to \autoref{subsec:related_work_online_trajectory_authoring}, our thesis shares the online programming setting and the use of spatial input, but we focus on a more high-level, intent-based, use case-driven approach, where the user does not need to specify the full sequence of waypoints or tool actions.

\section{Multimodal intent: combining gestures with language}

Beyond AR-only interaction, there is growing interest in combining gestures (e.g.\ pointing) with natural language to express intent at a higher semantic level than explicit waypoints. Lin et al.~\cite{Lin2023} propose \emph{GIRAF}, a framework that combines gesture recognition with large language model-based reasoning to interpret multimodal instructions in a context-dependent manner, and demonstrate it on tabletop manipulation tasks. Conceptually, this direction is promising because it can reduce the reliance on explicit GUI-based interfaces and instead enables direct natural language communication with the system.

That being said, the presented instantiation of \emph{GIRAF} is evaluated in a narrowly scoped tabletop setting and relies on a collection of task- and domain-specific perception, grounding, and reasoning components. Extending the approach to other tasks or environments would require substantial additional development. Moreover, \emph{GIRAF} primarily demonstrates \emph{interactive task assistance} from instructions, rather than authoring reusable robot programs in industrial workflows. In contrast, our thesis focuses on an extensible system for \emph{programming} task families through reusable sensing and authoring capabilities. Nevertheless, multimodal intent understanding (e.g., gesture and speech) is a promising direction for future work to further reduce friction in robot programming, and we discuss it further in \autoref{chapter:ch6_discussion}.

\section{Positioning of this thesis}

In summary, existing work shows multiple viable ways to reduce programming friction: AR-assisted online waypoint/path authoring with validation \cite{Ong2020}, rule-oriented online authoring in AR using trigger-action abstractions \cite{Ikeda2024}, and offline AR workflows that reduce robot downtime via transfer and export \cite{Muller2024}. Hand-tracking-based waypoint authoring is also promising for reducing friction, but it may require additional work to achieve the accuracy required for precise pose specification \cite{Vysocky2023}. Multimodal approaches such as \emph{GIRAF} show promise for enabling more natural communication of intent, but they remain early-stage and task-scoped \cite{Lin2023}.

Compared to these approaches, our thesis targets an extensible, task-oriented online spatial authoring system that attempts to avoid the need for users to specify low-level motion details by \emph{combining} spatial intent with sensed scene context, while reusing common sensing, authoring, and execution pipelines. This survey therefore establishes the relevant design space, while \autoref{chapter:ch2_analysis} turns that landscape into concrete requirements, comparison criteria, and the selected system direction.
