\chapter{User Workflow} \label{chapter:ch5_user_workflow}

This chapter describes the operator workflow of the prototype via the web frontend: how to start the system, instantiate and wire the module graph, activate and calibrate the sensing pipeline, and author and execute commands for the implemented use cases.
Build and installation steps for the submitted code snapshot are described in \autoref{appendix:source_code}, and the required physical artifacts (Charuco board, pen, fiducial box registry/CAD, and CBun sources) are summarized in \autoref{appendix:artifacts}.
The UI actions referenced below correspond to the orchestration behavior described in \autoref{sec:ch3_ui_orchestration}, and each workflow step can be mapped back to the instantiated configuration in \autoref{fig:runtime_full}.

\xxx{Make it a true “how to use it” chapter:}

\xxx{1. System setup workflow (instantiate + activate; can be shorter if mapping auto-selects).}

\xxx{2. User story: pick-and-place (end-to-end, with screenshots).}

\xxx{3. User story: weld (end-to-end, with screenshots).}

\xxx{4. Optional: persistence “save/load your setup” and “save/load your program” as small subflows referencing \autoref{sec:ch3_persistence_semantics}.}

\xxx{This chapter heavily references:}

\xxx{\autoref{sec:ch3_ui_orchestration} for UI screens and actions,}

\xxx{\autoref{fig:runtime_full} for “what modules are involved when you press X”.}




5.1 Preconditions and environment

\minihead{Physical prerequisites.}
The workflow assumes the following physical setup is available (see \autoref{sec:ch4_physical_setup} for context and \autoref{appendix:artifacts} for delivered files):
\begin{itemize}
    \item Kassow robot with the stereo camera rigidly mounted (mount CAD in \autoref{appendix:camera_mount_cad}).
    \item Printed Charuco board (\autoref{appendix:charuco}) placed rigidly in the workspace.
    \item At least one fiducial-tagged box object present in the workspace; the object model corresponds to the registry in \autoref{appendix:scene_registry} (example CAD in \autoref{appendix:scene_cad}).
    \item The pen is assembled, flashed, and charged; it advertises over BLE as \type{"3D Pen"} (see \autoref{appendix:pen_artifacts}).
\end{itemize}

\minihead{Software prerequisites.}
The core runtime and modules must be built and installed into the expected runtime layout (\path{binaries/}) as described in \autoref{appendix:source_code}.
This chapter assumes the system is run from the installed layout (modules under \path{binaries/modules/}, module data under \path{binaries/data/}).

\minihead{Robot-side prerequisites.}
The robot controller must have the Aergo CBun adapter installed and enabled (\path{robot/aergo_interface_cbun/}); the PC running the thesis runtime must have IP connectivity to the robot controller and the adapter’s TCP endpoint (IP/port).
CBun build notes and artifact locations are documented in \autoref{appendix:cbun}.

\minihead{Reproducibility boundary.}
End-to-end execution requires either a physical Kassow robot or a compatible offline environment (REMII). Access to REMII may be restricted; readers without access can still build and inspect the runtime and modules, but cannot validate robot execution.

5.2 System startup

\xxx{Start core + UI; confirm auto-created UI module exists (if applicable).}

\xxx{Load the intended configuration (either by building graph manually or by loading a saved \type{.aergo} setup).}

\xxx{If the Aergo CBun adapter is not yet installed on the robot controller: build it from \path{robot/aergo_interface_cbun/} (Dev Container workflow) and deploy the produced \path{build/aergo_adapter.cbun} to the controller.
Build/deploy artifact locations are documented in \autoref{appendix:cbun}; document only the robot-side installation/enabling steps here.}

5.3 Graph instantiation and wiring

\xxx{Show AddModuleUi for each module with “AUTO\_ALL to UI” assumption.}

\xxx{If mapping auto-selects because only one producer exists, say so explicitly and show only one full example; for the rest use a short “repeat similarly” pattern.}

\xxx{Reference \autoref{fig:runtime_full} repeatedly: “this wiring corresponds to edge X”.}

5.4 Activation and calibration workflow

\xxx{Activate robot module: set IP/port, verify connected.}

\xxx{Activate stereo camera module: select device, verify preview.}

\xxx{Calibration: explain Charuco placement + moving robot through poses conceptually; describe what the UI shows during calibration (progress, success/failure).}
The required Charuco target and its exact dimensions are provided in \autoref{appendix:charuco}.

Since no robot: \xxx{Use a “conceptual step list” + 1 diagram/screenshot from earlier logs if you have it; otherwise keep textual. Avoid claiming actual measured accuracy.}

5.5 User story 1: Pick-and-place

\xxx{Scene scan step.}
\xxx{Ensure the scene registry JSON is present in the module data directory (see \autoref{appendix:scene_registry}); then trigger a scene refresh in the UI and verify the detected object list.}

\xxx{Pick pose capture; place pose capture; confirm command; show preview overlay.}

\xxx{Run; expected motion sequence; show completion states.}

\xxx{Discuss what happens if object not found / tracking lost.}

5.6 User story 2: Weld

\xxx{Scene scan step; draw trajectory; confirm; preview overlay.}
\xxx{Scene detection uses the same registry mechanism as \autoref{appendix:scene_registry}; ensure the expected object types are present before drawing the trajectory.}

\xxx{Run; describe approach/start/end/depart structure.}

\xxx{Discuss matching failures and how user can adjust trajectory.}

5.7 Persistence workflows

\xxx{Saving/loading \type{.aergo} and what it restores (graph + activation params).}

\xxx{Saving/loading \type{.paergo} and the “missing use case id fails safely” behavior.}

\xxx{Out-of-sync commands: show “re-confirm required” behavior.}