\chapter{Discussion} \label{chapter:ch6_discussion}

This chapter returns to the question posed in the introduction: whether frequently changing collaborative-robot tasks can be programmed with less friction by specifying spatial intent directly in the real workspace. The aim here is not to restate the implementation chapter, but to interpret what the built prototype actually demonstrates, where its results are strongest, and where the boundary of those results lies.

\section{What the thesis built}

The main result of this thesis is an end-to-end prototype system for \emph{in-workspace spatial authoring} of selected robot tasks. The system combines workspace sensing, spatial input, task-specific interpretation, and robot execution behind a shared modular runtime with explicit interfaces. In practical terms, this means that the thesis did not only explore a tracked input device or a single interaction technique. It produced a complete workflow that covers module-graph construction, activation, calibration, authoring, confirmation, execution, visualization, and persistence, as described in \autoref{sec:ch3_ui_orchestration}, \autoref{sec:ch3_usecase_tree}, and \autoref{chapter:ch5_user_workflow}.

Taken against the goals stated in the introduction, the thesis achieves three results in a scoped but defensible sense. With respect to \goalref{goal:1}, it realizes direct spatial authoring in the real workspace by allowing the user to enter poses and trajectories with a tracked pen, combine those inputs with sensed workspace context, and execute the resulting task through a shared runtime. With respect to \goalref{goal:2}, it provides an extensible architecture in which task-oriented use cases are implemented as plugins on top of the same acquisition and execution infrastructure. With respect to \goalref{goal:3}, it demonstrates the approach on a real Kassow robot using two representative task families, pick-and-place and seam welding, and uses those use cases to discuss both the benefits and the present limitations of the approach.

This connection to the introduction matters, because the thesis result is broader than the two example tasks alone. The system described in \autoref{chapter:ch3_system_design} separates the generic runtime contract from the application-level interfaces and then separates both from the concrete modules that implement sensing, scene detection, robot execution, and use-case logic. This makes the strongest conceptual result of the thesis the \emph{combination} of shared runtime, explicit interfaces, reusable use-case boundary, and complete operator workflow. The tracked pen is one part of that result, but it is not the whole contribution.

The use-case boundary is especially important here. The use case interface and the \type{UsecaseTree} helper described in \autoref{sec:ch3_usecase_wrapper} and \autoref{sec:ch3_usecase_tree} turn the underlying runtime into a concrete programming model: the user can discover available use cases, create commands, confirm them into executable representations, run them sequentially, and save or load the authored program. This is why the system should be read as a reusable prototype framework for task-oriented spatial authoring rather than as two disconnected demonstrations that happen to share some code.

The implementation chapter supports the same reading. The two evaluated use cases reuse the same sensing and execution backbone, and the motion-primitive use cases in \autoref{sec:ch4_usecases_primitives} show that the runtime is not hard-coded only to pick-and-place and welding. Those motion primitives are not the main practical outcome of the thesis, but they do show that the same use-case boundary can host both task-specific procedures and lower-level command forms. That increases confidence that the runtime is a genuine system contribution and not merely a custom scaffold built around one narrow demo.

\section{What the results mean}

The practical idea explored by the thesis is not to replace all robot programming, but to shift a certain class of work away from repeated low-level editing and toward fast re-authoring inside a prepared task-specific interface. In many real deployments, the initial integration effort goes into defining the task, calibrating the environment, and connecting sensing and execution. The presented system keeps that one-time integration effort in the shared runtime and the prepared use case, so that the operator can later redefine only the parts of the task that vary from one task instance to the next.

This is why the modular architecture matters as a result in its own right. Chapters~2 and~3 argue for stable interfaces between sensing, UI, robot backend, and use-case logic, and the implemented system shows that this separation is practical. The two main use cases are not simply two examples of ``pen programming''; they are two different task interpretations built on the same infrastructure. That is a stronger thesis result than showing one carefully tuned workflow for one use case, because it demonstrates that the core system can support different task structures while keeping the task-specific logic local to the use case module.

This also sharpens the thesis' position relative to the related work reviewed in \autoref{chapter:ch1_related_work}. The result is not primarily a waypoint- or path-authoring system, not primarily a rule-based AR authoring system, and not primarily an offline transfer workflow. Instead, the distinguishing contribution of this thesis is a task-oriented runtime in which spatial intent and scene context are combined inside reusable use cases that share the same sensing, orchestration, and execution infrastructure.

At the same time, the two use cases do not carry the same weight. Pick-and-place is important because it shows that the system is not limited to replaying fixed trajectories. The authored command stores a selected object identity and a pick pose relative to that object, and at execution time the system re-reads the scene and reconstructs the world-frame pick pose from the current object pose. This is a meaningful architectural result: it shows that authored intent and current scene state can be combined at runtime instead of being frozen at confirmation time. In that sense, pick-and-place demonstrates that the runtime can support scene-dependent execution rather than only fixed replay.

However, pick-and-place is not the strongest practical target for the current prototype. The present scene model is intentionally narrow, based on registry-defined fiducial boxes, and the use case depends heavily on that narrow model. More broadly, many industrial handling tasks are highly specific and are usually integrated together with dedicated grippers, conveyors, fixtures, and vision systems. Such applications often require a tighter coupling between tooling, sensing, and factory process than the current prototype provides. The thesis therefore supports a more modest conclusion: pick-and-place is valuable primarily because it proves that the architecture can combine authored spatial intent with runtime scene information, not because it already solves generic industrial material handling.

Seam welding is the clearest practical fit for the thesis approach. The task is naturally trajectory-centric, the user intent is spatially meaningful, and the use case can infer most of the internal robot procedure from the combination of authored trajectory and sensed scene geometry. In the implemented workflow, the user scans the scene, draws the intended weld, confirms the interpreted result, and then executes an approach--weld--depart sequence. That workflow is simple enough to be direct, yet structured enough to show how spatial authoring can reduce the amount of explicit robot programming exposed to the operator.

More importantly, welding matches the motivating scenario of the thesis better than pick-and-place does. The core motivation in the introduction is not only that repetitive tasks exist, but that some tasks change often enough that repeated low-level reprogramming becomes a real cost. Welding-like tasks, especially in small-batch or frequently re-taught settings, fit this pattern well. The value of the approach is therefore strongest where the physical operation is repetitive enough that robot execution is worthwhile, but the concrete welds change often enough that repeatedly editing a conventional robot program becomes inefficient. In that setting, the operator can indicate the weld in context and let the system infer the rest of the procedure, which is exactly the interaction pattern the thesis set out to explore.

The discussion also clarifies an important trade-off in the use-case model itself. A highly specialized use case can be prepared for one narrow recurring workflow and expose only the spatial inputs that actually vary. That gives the most direct operator experience, but it also limits reuse across deployments. A more generic use case can be reused more broadly, yet it must expose more parameters and perform a less direct mapping from intent to execution. The thesis does not resolve the best balance between these two extremes, but it does show that both are compatible with the same underlying runtime and interface design.

This leads to a final interpretive point. The thesis is strongest when the system is read as a middle ground between two extremes: on one side, fixed industrial integration for one stable task; on the other side, a general robot programming environment. The presented approach does not aim to subsume the second extreme. Its main value is to give the operator a fast way to redefine a known task family inside a prepared use case, especially where repeated low-level reprogramming would otherwise dominate the cost of using the robot.

\section{Where the approach works well}

Several positive results are supported clearly enough by the rest of the thesis to be stated directly. First, the task-oriented authoring model appears to be a good compromise between low-level robot programming and highly abstract intent systems. The operator does not need to specify every internal motion, yet the interaction remains grounded in concrete world-space input and a well-defined executable procedure. This balance is especially strong in use cases such as welding, where the system can infer substantial internal structure from a trajectory authored directly in context.

Second, the modular architecture is not only an implementation convenience, but one of the stronger contributions of the thesis. Chapters~2 and~3 argue that sensing, interpretation, UI, visualization, and robot execution should be exchangeable behind stable contracts, and the implemented prototype confirms that this separation is workable in practice. Input modality, scene detection, camera configuration, UI frontend, and robot backend are all treated as modules behind explicit interfaces rather than as assumptions baked into one monolithic application.

Third, the use-case boundary proved to be an effective abstraction layer. The frontend communicates with the wrapper protocol rather than with use-case-specific APIs, while the use cases themselves reuse the same sensing and execution infrastructure. This is a meaningful software-design result: it localizes task-specific logic while preserving a shared runtime and a shared operator workflow. The motion-primitive examples reinforce the same point, because they show that the use-case boundary can host both narrowly task-oriented behavior and lower-level command forms.

Finally, the end-to-end workflow itself should be counted as a successful result. The prototype is able to carry the user from setup and activation through authoring, confirmation, execution, visualization, and persistence inside one system. Even where the implementation remains rough, this makes the thesis more than an interaction sketch. It demonstrates that in-workspace spatial authoring can be embedded into a reusable runtime and exercised on a real robot in a way that is coherent from initial setup to final execution.

\section{Current limitations}

The thesis result should therefore be read as a prototype-level result with a clearly bounded scope. Some of those limits follow from the thesis concept itself, while others arise from the current implementation. Put differently, some boundaries are inherent to the chosen task-oriented programming model, while others reflect the present maturity of the prototype used to evaluate it.

\minihead{Scope and programming model.}
The first boundary is conceptual. As stated already in the introduction, the thesis does not target a full robot programming language. It does not address rich branching logic, loops, exception recovery, or more general sensor-driven program structure. The programming model exposed by the current frontend is intentionally linear and use-case-oriented. This is not a flaw accidentally discovered late in the project; it is the scope that made it possible to focus on whether spatial intent input can simplify authoring for selected task families. The consequence is that the present system is best understood as a framework for prepared task classes, not as a general replacement for established robot-programming environments.

\minihead{Sensing and calibration.}
The second boundary concerns sensing robustness. Pen tracking is marker-based and therefore sensitive to visibility, motion blur, lighting conditions, and the quality of corner detection. The current pen implementation does not fuse IMU data even though the device provides it, and the camera-pose calibration assumes a rigid robot-mounted stereo rig whose calibration remains valid only while the mount is mechanically unchanged. These limits do not invalidate the prototype demonstrations, but they do mean that the sensing stack has not been characterized as a robust deployment solution for broader factory conditions.

\minihead{Scene understanding.}
The strongest practical limitation of the current prototype is scene understanding. The implemented scene-detection interface operates on registry-defined fiducial boxes, returns only object IDs and poses, and processes one stereo pair per \type{READ\_SCENE} request without temporal stabilization or confidence outputs. This was sufficient to validate the architectural flow from scene sensing to task interpretation, but it sharply constrains how far the demonstrated workflows can be transferred beyond the present examples. In particular, the current system does not provide a richer object model, uncertainty handling, or broader scene semantics of the kind that many real handling or fabrication tasks would require.

\minihead{Execution and tooling.}
The robot execution layer is also intentionally incomplete. The thesis implements a shared robot motion interface and demonstrates that use cases can execute real robot motion through it, which is an important result. At the same time, generalized tool control is not yet standardized in the interface, and the evaluated use cases simulate gripper or welding actions by fixed delays. No collision checking is performed in the current use cases, and the prototype does not support simulated execution of authored programs. For fixed tasks, export or compilation into a lower-level robot-side representation may be possible in the future, but this is not implemented in the present system.

\minihead{UI maturity and portability.}
The frontend should likewise be read as a prototype control-plane interface rather than as a finished production UI. It already covers the full workflow from setup through execution and persistence, which is one of the strengths of the thesis, but some of its behavior still reflects prototype maturity. For example, the operator may need to explicitly reload the visualization when stale artifacts remain visible, and reloading a saved setup can fail or restore only part of the graph if hardware modules cannot reactivate. Similarly, the overall architecture is largely platform-neutral, but the concrete demonstrated workflow remains partly platform-bound because camera acquisition is currently Windows-only and the BLE discovery behavior differs between Windows and Linux in the present pen implementation.

\minihead{Applicability and safety.}
These technical limits also define the current applicability boundary. The weld use case assumes that the workpiece configuration remains unchanged after confirmation, while pick-and-place assumes that the scene can be refreshed at execution time and that the expected object can still be matched by its stored identity. More generally, the prototype assumes a clear workspace and does not perform collision-aware planning. The correct interpretation is therefore that the thesis demonstrates a research prototype for task-oriented spatial authoring, not a production-ready robot programming platform.

\section{Evaluation boundary}

The evaluation achieved by the thesis is qualitative and feasibility-oriented. This deserves to be stated separately, because the discussion chapter functions as the qualitative evaluation of the work and should make clear what kind of claims the thesis does and does not support.

The system was exercised on representative use cases, with most development and testing performed in the simulator and with limited validation on physical hardware. The thesis does not include a formal user study, a controlled comparison against teach pendant or offline programming workflows, or a quantitative characterization of tracking accuracy, repeatability, or robustness. It therefore does not support strong claims such as measured productivity gains, broad industrial readiness, or general superiority over existing programming methods.

There is also an inevitable operator and deployment boundary. The thesis reflects prototype development and validation in a narrow setting rather than multi-operator evaluation across varied industrial environments. The demonstrated tasks and hardware are representative enough to support the architectural and workflow claims made here, but they are not broad enough to establish general performance across task classes, sensing conditions, or deployment constraints.

Taken together, these limits define the level of validation achieved here. The thesis establishes feasibility for a modular system of task-oriented spatial authoring, shows that the idea can be exercised on substantially different task structures, and helps identify the class of tasks for which it appears most promising: tasks narrow enough to fit a prepared use case, but variable enough that repeated low-level reprogramming becomes the dominant cost. The next chapter turns from that interpretation of the present result to the directions in which the work can be extended.
