\chapter{Conclusion and Future Work} \label{chapter:ch7_conclusion}

\section{Conclusion}

This thesis investigated whether collaborative-robot tasks that change frequently can be programmed more directly by specifying spatial intent in the real workspace. Within prototype scope, the answer is positive for selected task-oriented workflows. Spatial intent in the real workspace can be combined with shared sensing, interpretation, and execution infrastructure to author and run prepared use cases without exposing the operator to a full robot programming environment.

The main contribution is therefore not only a tracked input device or two isolated demonstrations, but a modular runtime with explicit interfaces between sensing, authoring, interpretation, visualization, and execution. On top of that runtime, the thesis implements task-oriented use cases that reuse the same acquisition and execution infrastructure and exposes a concrete programming model in which use cases are discovered, confirmed into executable commands, executed sequentially, and persisted as user-authored programs. In this respect, the thesis fulfills the three goals stated in the introduction: it realizes direct spatial authoring in prototype form, it provides an extensible plugin-based architecture for use cases, and it demonstrates the approach on a real Kassow robot with representative workflows for pick-and-place and seam welding.

The strongest conceptual result is the combination of shared runtime, explicit interfaces, reusable use-case boundary, and end-to-end workflow. The strongest practical result is narrower: the approach appears best suited to trajectory-centric task families, with seam welding providing the clearest current fit. Pick-and-place remains an important architectural demonstration because it shows that authored intent can be combined with refreshed scene information at execution time, but it also depends more strongly on the present scene model and on prototype-level robustness.

These results define a technically coherent direction for task-oriented spatial authoring rather than a finished industrial programming platform. A reusable system and workflow of this kind can be built, integrated, and demonstrated on real hardware. At the same time, most development and testing remained in simulation, physical validation was limited, and stronger sensing, richer scene semantics, safer execution, and more formal evaluation are still required. The thesis demonstrates feasibility and architectural coherence, not measured superiority over existing workflows and not production readiness.

\section{Near-Term Future Work}

\minihead{Evaluation and validation.}
The most important next step is formal evaluation. The current thesis establishes feasibility, architectural coherence, and a plausible workflow, but it does not yet measure user benefit or operational robustness in a controlled way. A natural continuation would therefore combine a structured user study with stronger quantitative characterization of tracking quality, scene estimation, repeatability, workflow timing, and real-robot reliability. This would make it possible to answer more rigorously when spatial authoring reduces programming effort, for which task classes it is most useful, and how it compares with teach pendant or offline programming workflows.

\minihead{Sensing and scene pipeline.}
Another immediate direction is to strengthen the sensing stack, which currently limits the prototype most directly. This includes making better use of the sensing already present on the pen side, especially IMU data, improving calibration robustness and diagnostics, and moving beyond the present fiducial-box scene representation toward a richer and more stable scene estimate. In the current prototype, scene detection is intentionally simple and serves mainly to validate the architectural flow. A stronger research platform would benefit from temporal filtering, better uncertainty handling, and a scene representation that is less tightly coupled to the current scene abstraction used in the demonstrated workflows.

\minihead{Execution layer.}
A second engineering direction is to mature the execution backend from a minimal motion interface into a more complete task-execution substrate. The most important steps here are a generalized tool-control abstraction, clearer capability reporting at the robot-interface boundary, explicit motion validation, collision checking, safer planning behavior for authored tasks, and simulated execution support for authored programs. It would also be valuable to explore when a use case should be executed dynamically through the thesis runtime and when it could instead be exported or compiled into a lower-level robot-side representation, especially for tasks that do not require runtime scene adaptation.

\minihead{Prototype completion.}
The remaining near-term work is mainly completion work required to make the present prototype a more robust platform for further experiments. This includes completing the cross-platform boundary of the demonstrated workflow, improving persistence and orchestration robustness, and reducing the amount of operator recovery needed in cases such as visualization resynchronization or partial setup reload. These tasks are less conceptually central than sensing, execution, and evaluation, but they follow directly from the current prototype limitations and would materially improve the usability of the system as a basis for continued research.

\section{Longer-Term Directions}

Beyond prototype completion, the thesis opens several broader research directions. The first is richer interaction. The current system uses a tracked pen because it offered the most practical balance of stability, cost, and implementation effort under the thesis constraints, but the architecture is not tied to that choice. Chapter~2 already motivates alternative 6DoF inputs such as hand tracking and VR or AR controllers, and Chapter~1 points to multimodal intent as a promising longer-term direction. These alternatives could be explored within the same modular structure while keeping the core idea of spatially grounded, task-oriented authoring.

Another direction is integration with existing robot-programming workflows rather than keeping spatial authoring as a separate frontend. One possibility is to embed the same authoring ideas more directly into a teach pendant or controller workflow, so that spatial input becomes one tool inside a broader programming environment instead of a separate prototype application. This would also clarify how far the current linear use-case program model should be extended, and where spatial authoring is best treated as a complete task-oriented workflow rather than as an authoring aid inside more conventional robot-programming practice.

Finally, the system should be exercised on a broader set of task families. Chapter~2 already treats additional use cases such as painting and lower-level motion primitives as plausibility checks for the architecture, but the present thesis still evaluates only two main application families. Extending the work toward richer welding scenarios, dispensing and surface-processing tasks, or other human-in-the-loop workflows would help clarify where task-oriented spatial authoring is genuinely advantageous between fixed one-purpose integration and general robot programming. That broader exploration would also provide a stronger basis for both technical refinement and future evaluation.
