\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}

Collaborative robots are attractive for automating tasks that are repetitive but frequently changing, such as small batch production or tasks requiring frequent changeovers. In such cases, the bottleneck is often not the execution time itself, but the time and effort required to re-create robot programs for each new task.

Robot programming is both time-consuming and skill-intensive. This creates friction whenever tasks need to be re-taught. This thesis explores whether this friction can be reduced by changing how users specify robot tasks.

\section*{Motivation}
\addcontentsline{toc}{section}{Motivation}

Industrial robots are usually programmed by using one of the following methods \citep{DevicVidakovicZivkovic24}:
\begin{itemize}
    \item \emph{Offline programming}: The user creates the robot program in a simulated environment, often based on a CAD model of the workspace. This approach allows programming without the physical presence of the robot, therefore avoiding downtime and allowing testing in a simulated environment. However, it requires accurate models of the robot and the environment, and it can be complex to set up and use effectively.
    \item \emph{Teach pendant programming}: The user enters robot motions and commands step-by-step using a handheld device (the teach pendant). For collaborative robots, this is often complemented by \emph{hand-guiding} \footnote{Terminology varies across sources. In this thesis, \emph{hand-guiding} refers to physically back-driving the robot to teach poses, while \emph{lead-through} is reserved for continuous path recording.}, where the user can physically back-drive the robot to teach poses. While this can make pose specification more intuitive, it remains time-consuming to enter precise poses and trajectories and requires the user to be physically present at the robot, which can lead to downtime during programming.
\end{itemize}

Both of these methods are widely used and capable of producing effective robot programs, but they can become friction-heavy when tasks change frequently. Robot programs usually consist of a sequence of move commands interleaved with tool actions and logic. This brings two main sources of friction:
\begin{itemize}
    \item \emph{Pose definition is tedious}: Specifying precise poses and trajectories through a teach pendant or offline programming can be time-consuming and unintuitive, especially for users without extensive programming experience. This can lead to long iteration loops of editing the program, testing it on the robot, and adjusting it until the desired behavior is achieved.
    \item \emph{Expertise and setup overhead}: Both teach pendant and offline programming require a certain level of expertise in robot programming concepts, and offline programming also requires accurate models and complex setup. This can be a significant barrier for users who are not familiar with programming concepts, and it can lead to reliance on trained personnel or system integrators for programming and deployment.
\end{itemize}

Collaborative robots often partially mitigate the pose definition friction with hand-guiding the robot \citep{KassowRobotsManuals}. However, hand-guiding may not be suitable for all tasks and it does not solve the expertise and setup overhead.

This motivates an alternative approach: instead of programming motions and tool actions directly, let the user specify spatial intent (e.g., a pose or a coarse trajectory) in the real workspace and let task-specific use cases transform this intent into executable robot motions and tool commands using scene context. This approach, which we call \emph{in-workspace spatial authoring}, allows users to focus on the task logic and spatial intent rather than the low-level details of robot programming, making it more accessible to a wider range of users, while also potentially speeding up the programming process for experienced users.
