%%% Please fill in basic information on your thesis, which will be automatically
%%% inserted at the right places. You need to replace \xxx{...} by real data.

% Type of your thesis:
%	"bc" for Bachelor's
%	"mgr" for Master's
%	"phd" for PhD
%	"rig" for rigorosum
\def\ThesisType{mgr}

% Language of your study programme:
%	"cs" for Czech
%	"en" for English
\def\StudyLanguage{cs}

% Thesis title in English (exactly as in the official assignment)
% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{Defining robot trajectory using augmented reality}

% Author of the thesis (you)
\def\ThesisAuthor{Jakub Medek}

% Year when the thesis is submitted
\def\YearSubmitted{2026}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% see https://www.mff.cuni.cz/en/faculty/organizational-structure,
% or a full name of a department outside MFF)
\def\Department{Department of Theoretical Computer Science and Mathematical Logic}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

% Thesis supervisor: name, surname and titles
\def\Supervisor{RNDr. David Obdržálek, Ph.D.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Department of Theoretical Computer Science and Mathematical Logic}

% Study programme (does not apply to rigorosum theses)
\def\StudyProgramme{Computer Science - Artificial Intelligence}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, who provided you with tea and pizza, etc.)
\def\Dedication{%
I would like to thank my supervisor, RNDr. David Obdržálek, Ph.D., for his guidance and help during the creation of this thesis.
I would also like to thank my family for their continued support and encouragement throughout my studies.
}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
This thesis explores spatial authoring of cooperative robot programs in the real workspace as an alternative to teach pendant and lead-through methods. Instead of programming motions directly, the user provides spatial intent (e.g., a pose, or a coarse trajectory) via AR-inspired inputs, and task-specific use cases transform this input into executable robot motion and tool commands using scene context.\newline
A modular runtime is developed to support task-oriented programming use cases as plugins that reuse the same world-space acquisition and robot execution pipelines, while allowing components such as sensing, tracking, and scene estimation to be replaced.\newline
The approach is implemented on a Kassow Robots collaborative robot with camera-based workspace referencing, a tracked 3D pen, and a tablet-oriented web UI. Representative use cases --- pick-and-place and seam welding --- are developed and tested on the real robot, together with basic motion primitives.\newline
Experiments evaluate tracking stability, end-to-end authoring time, and failure modes. Results demonstrate that spatial intent input can significantly reduce setup effort for frequently redefined, trajectory-centric tasks such as welding, and they identify practical limitations driven by sensing accuracy and tool integration.
}

% 3 to 5 keywords (recommended) separated by \sep
% Keywords are useful for indexing and searching for the theses by topic.
\def\ThesisKeywords{%
spatial authoring\sep
collaborative robots\sep
augmented reality\sep
robot programming\sep
modular software architecture%
}

% If any of your metadata strings contains TeX macros, you need to provide
% a plain-text version for use in XMP metadata embedded in the output PDF file.
% If you are not sure, check the generated thesis.xmpdata file.
\def\ThesisAuthorXMP{\ThesisAuthor}
\def\ThesisTitleXMP{\ThesisTitle}
\def\ThesisKeywordsXMP{\ThesisKeywords}
\def\AbstractXMP{\Abstract}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  % uncomment to decrease font size

% If you are studing in a Czech programme, you also need to provide metadata in Czech:
% (in English programmes, this is not used anywhere)

\def\ThesisTitleCS{Zadávání trajektorie robota pomocí rozšířené reality}
\def\DepartmentCS{Katedra teoretické informatiky a matematické logiky}
\def\DeptTypeCS{Katedra}
\def\SupervisorsDepartmentCS{Katedra teoretické informatiky a matematické logiky}
\def\StudyProgrammeCS{Informatika - Umělá inteligence}

\def\ThesisKeywordsCS{%
prostorové programování\sep
kolaborativní roboti\sep
rozšířená realita\sep
programování robotů\sep
modulární softwarová architektura%
}

\def\AbstractCS{%
Tato práce zkoumá prostorové zadávání programů pro~kolaborativní roboty přímo v~reálném pracovním prostoru jako alternativu k~programování pomocí teach pendantu a k~metodám lead-through. Uživatel nezadává pohyby robota přímo; místo toho poskytuje prostorový záměr (např. požadovanou polohu nebo hrubý průběh trajektorie) pomocí vstupů inspirovaných rozšířenou realitou. Úlohově orientované moduly následně tento záměr s~využitím kontextu scény převádějí na~vykonatelné pohyby robota a příkazy pro~nástroje připevněné k~robotu (např. otevřít/uzavřít gripper).\newline
V~rámci práce byl vytvořen modulární běhový systém, který umožňuje implementovat a rozšiřovat programovací use-casy formou pluginů. Ty sdílejí rozhraní pro~získávání dat ve~světových souřadnicích a pro~vykonání programu na~robotu, přičemž jednotlivé části (snímání a sledování polohy, detekce scény) lze vyměnit za přesnější nebo méně experimentální řešení.\newline
Navržený přístup je realizován na~kolaborativním robotu Kassow Robots s~detekcí scény a sledováním 3D~pera pomocí stereo kamerového systému a webovým rozhraním optimalizovaným pro~použití na~tabletu. Byly implementovány a na~reálném robotu otestovány reprezentativní úlohy pick-and-place a seam welding, doplněné o~základní pohybové primitivy.\newline
Experimenty hodnotí stabilitu sledování, čas potřebný k~vytvoření programu od~začátku do~spuštění a typické režimy selhání. Výsledky ukazují, že zadávání prostorového záměru může výrazně snížit čas i~náročnost přípravy u~často upravovaných, na~trajektorii založených úloh, jako~je svařování, a zároveň odhalují praktická omezení daná přesností snímání a integrací nástrojů.
}
